Wireless Automobile Detection, License Plate Processing, and Data Availability Network Proposal

### 2014-03-17 proposal/proposal.tex
Edited deployment, summary, and acquisition sections

The server will be running a web application with an HTTP interface that will both receive the data from the base station and route it to the correct part of the application (database, image processor, etc.). This interface will be the method of communication between the server and the base station. When data is received from the base station, the application will process it, store it in a database, and display it appropriately on the website. The following sections describe the architecture and function of each of the aforementioned subsystems in detail. The detection subsystem is broken up into two parts: Automobile Detection (Section 2.2) and License Plate Image Acquisition (Section 2.3). The server subsystem is also broken into two parts: Server Processing (Section 2.5) and Web Application (Section 2.6).

\subsection{Automobile Detection}
The automobile detection subsystem provides a means for detecting ingress and egress of vehicles from a parking lot. This subsystem is to be placed on the side of the road next to each parking lot entrance and exit. Automobiles will be detected using a magnetometer which perceives the induced change in the local magnetic field as the metallic structure of the vehicle passes by, as described in \cite{stillwell2013}. The automobile detection subsystem will be based around the commercially available Arduino Fio 16-bit  platform which utilizes the ubiquitous Atmel ATMEGA328p microcontroller. When an automobile is detected by the Arduino Fio, a packet is sent through the onboard XBee Pro radio to the central base station (described in Section 2.4). The packet will  contain the direction of the passing vehicle. At the time of detection,  a hardware interrupt is also sent to the license plate recognition system to awaken it from sleep.

@@

\subsection{License Plate Image Acquisition}
Kevin
From Brandon: I called out your section in my section to talk about the determination of cropping and image manipulation on the server versus the Acquisition Raspberry Pi.

\subsection{Central Base Station}
The central base station will be the point of contact between the sensor network and the server. The data gathered by the Arduino Fio and Raspberry Pi sensors will be routed to the base station via XBee Pro radios. The base station, which will itself be a Raspberry Pi, will forward the data to the ACMx server using HTTP requests. Forwarding data is the main function of the base station. The base station will run the Raspbian operating system \cite{raspbian}, a lightweight version of Linux. Because Raspbian is optimized for the Raspberry Pi, It should be easy to write some shell scripts that will take care of forwarding the data via HTTP to the server. A second function of the base station will be to take periodic overhead photos of the parking lots to help reconcile the sensed data with the actual state of the lots. System administrators will be able to look at the photos to see if the number of cars reported in the lot by the sensors is correct, and fix any errors. See section 2.6 for more detailed information about the error correction functionality of the web application.

@@

Q-Free Intrada ALPR is a license plate recognition software that offers a C++ API, as well as a cloud-based service. Both the API and cloud-based service can be utilized from a Linux server. Q-Free's software is used in countries around the world for traffic management and toll collection. The wide-ranging geography of its use mean that Intrada ALPR is likely to be reliable and accurate for all types of license plates. Use of the Intrada suite is dependent on Q-Free granting an educational use license for this semester.

OpenALPR is another C++ library for use with both North American and European plates. This library relies upon two underlying technologies: OpenCV (an open-source computer vision library) and Tesseract OCR (an OCR engine being developed and maintained by Google). The Tesseract OCR Tool is self-sufficient, and relies on included training data. \cite{patel2012} goes into greater detail how Tesseract extracts data from images. The open-source nature of this project make it appealing as it can be immediately integrated without needing to obtain an educational license first. Additionally, the last updates were pushed to the source repository in January. While it is desirable for the software we consider to be relatively up-to-date and maintained, this repository was started only four months ago (November, 2013). Use of such a young library may create a less stable solution than what is practical to work with throughout the course of the semester.

@@

Upon upload of the image to the ACMx server, a script will be invoked to process the image. The ACMx server will then run redundancy measures to ensure the image uploaded does contain a license plate. Upon a successful identification, the server will invoke the selected ALPR strategy to extract characters from the license plate image. Cropping and image manipulation required by the selected ALPR strategy will be done before invoking this script. The amount of cropping done on the server and on the Acquisition Raspberry Pi is something to be determined throughout the course of this project. Section 2.3 goes into more detail about this subject.

@@

Once the ACMx server has a plain text representation of the license plate characters, they will be stored in a MySQL database, along with a timestamp and lot id, where they can be retrieved and processed by the Web Application subsystem described in Section 2.6.

\subsection{Web Application}
As mentioned in Section 1.2, the ACMx group has already created a web application that will display the data collected from the sensors in the system. The site is currently accessible online at \url{http://acmxlabs.org/parking/}. Since the current site is essentially just a shell with no data, one main goal of this project is to make the main pages of the site fully functional. Another main goal is to create an administrator side to the site, where data can be viewed and updated by site administrators. The main focus of this subsystem will be to provide interfaces for both general users and administrators that make the data we collect both useful and easy to access. Since the ACMx team already has some protocols in place for collecting and displaying the data from the magnetometer Arduino Fios, the ACMx team will continue development on this interface, and our group will focus on other parts of the application, as described below. The web application itself will be written with PHP on the back end and HTML5/CSS3/JavaScript on the front end. The following subsections describe the various pieces of the site that we aim to complete during this project.

@@

\subsubsection{Validating and Adjusting Lot Capacity Data}
We expect that our first deployment of this system may not be 100\% accurate at detecting all vehicles that enter and exit lots. Therefore, as mentioned in section 2.4, we will be collecting periodic overhead images from a camera mounted on the base station Raspberry Pi. These images will serve as a visual representation of how many cars are actually in the parking lot at the time the image was taken. Administrators should be able to periodically compare this image to the collected data from the sensor network to make sure the numbers being computed by the network accurately represent the number of cars in the lot. As previously mentioned, the deployment of this project will only be on two campus parking lots, and we may only have one overhead camera to start off with, although ideally we would have two - one for each lot. The web interface will allow administrators to pull up the image for any lot with an overhead camera and see the numbers for that lot right next to the image. If any discrepancies are found, the administrator should be able to overwrite the count in the database to reflect reality. This interface will be extremely useful in helping to calibrate the system. For example, if the sensors are consistently reporting less cars in the lot than the image shows, we can fine-tune the collection process to make sure fewer cars slip through unnoticed. Future work in this area will include an automated version of this process, where a computer program will compare the image to the sensed data and make any necessary corrections.

@@

\subsection{Deployment}
The system is planned to be deployed at the CTLM Upper and Lower parking lots at CSM. These parking lots were chosen because of their accessible location, unobstructed visibility from a campus structure, and wide use. Both parking lots have two entryways, each of which will be equipped with an automobile detection subsystem as shown in Figure \ref{fig:deployment}. All entryways also benefit from having a clear view of the sky for increased photovoltaic efficiency.

\begin{figure}
\begin{center}
\includegraphics[width=5in]{deployment}
\end{center}
\caption{Deployment Location}
\label{fig:deployment}
\end{figure}

The central base station, as described in Section 2.4, will be located on the second level of the CTLM building. Several potential locations exist which allow for a clear line-of-sight to all automobile detection subsystems, one of which is detailed in Figure \ref{fig:deployment}. These locations also provide internet access to the base station Raspberry Pi through ethernet ports.

\section{Summary}
A summary \cite{johnson2012} Talk about the motivation behind this project. Why it is important.

@@

>>>>

The server will be running a web application with an HTTP interface that will both receive the data from the base station and route it to the correct part of the application (database, image processor, etc.). This interface will be the method of communication between the server and the base station. When data is received from the base station, the application will process it, store it in a database, and display it appropriately on the website. The following sections describe the architecture and function of each of the aforementioned subsystems in detail. The detection subsystem is broken up into two parts: Automobile Detection (section 2.2) and License Plate Image Acquisition (section 2.3). The server subsystem is also broken into two parts: Server Processing (section 2.5) and Web Application (section 2.6).

\subsection{Automobile Detection}
The automobile detection subsystem provides a means for detecting ingress and egress of vehicles from a parking lot. This subsystem is to be placed on the side of the road next to each parking lot entrance and exit. Automobiles will be detected using a magnetometer which perceives the induced change in the local magnetic field as the metallic structure of the vehicle passes by, as described in \cite{stillwell2013}. The automobile detection subsystem will be based around the commercially available Arduino Fio 16-bit  platform which utilizes the ubiquitous Atmel ATMEGA328p microcontroller. When an automobile is detected by the Arduino Fio, a packet is sent through the onboard XBee Pro radio to the central base station (described in section 2.4). The packet will  contain the direction of the passing vehicle. At the time of detection,  a hardware interrupt is also sent to the license plate recognition system to awaken it from sleep.

@@

\subsection{License Plate Image Acquisition}
Once the Acquisition Raspberry Pi receives the interrupt from the vehicle detection Arduino Fio, it will begin the process of obtaining a picture of the vehicle's license plate. This procedure will be separated into two separate sections: image acquisition and license plate detection.

\subsubsection{Image Acquisition}
While the automobile detection subsystem correctly identifies when a vehicle enters the lot, the image acquisition subsystem will be responsible for determining when to capture a picture of that car in order to optimize the quality of the license plate image. The timing difference between when the Raspberry Pi receives the interrupt and when it actually captures the picture will initially be determined experimentally using the average speed at which cars enter the parking lot. As the project progresses, this timing will be fine-tuned in order to capture the optimal image every time. The goal of this subsystem is to only take one picture to capture the license plate, so as to minimize capture time and power consumption.

\subsubsection{License Plate Detection}
Once the image has been captured, the Raspberry Pi will then focus on detecting the license plate in the image and cropping the image to only include the license plate. This will be done in order to minimize the size of the transmission and the power required for that transmission. In order to detect the location of the license plate within the larger image, a detection algorithm needs to be applied.

The location of the license plates will be detected on the acquisition Raspberry Pi using the OpenCV (Open Source Computer Vision Library) API \cite{openCV}, which provides many different algorithms that can be used for image detection and shape recognition. This particular API was chosen for both its ease of use and widespread cross-compatibility; it has interfaces for development in C, C++, Python, Java, and MATLAB, and supports all major operating systems (including distributions of Linux such as Raspbian). Because OpenCV is used across the industry for image processing, we are confident that it will meet our needs.

Once the OpenCV software has detected the license plate, the image will then be cropped as needed until all that remains in the image is the license plate from the car that is entering the lot. That image will then be sent to the central base station via an XBee Pro radio transmission.

\subsection{Central Base Station}
The central base station will be the point of contact between the sensor network and the server. The data gathered by the Arduino Fio and Raspberry Pi sensors will be routed to the base station via XBee Pro radios. The base station, which will itself be a Raspberry Pi, will forward the data to the ACMx server using HTTP requests. Forwarding data is the main function of the base station. The base station will run the Raspbian operating system \cite{raspbian}, a lightweight version of Linux. Raspbian is optimized for the Raspberry Pi, so writing shell scripts that will take care of forwarding the data via HTTP to the server will be an easily achievable task. A second function of the base station will be to take periodic overhead photos of the parking lots to help reconcile the sensed data with the actual state of the lots. System administrators will be able to look at the photos to see if the number of cars reported in the lot by the sensors is correct, and fix any errors. See section 2.6 for more detailed information about the error correction functionality of the web application.

@@

Q-Free Intrada ALPR \cite{intrada2014} is a license plate recognition software that offers a C++ API, as well as a cloud-based service. Both the API and cloud-based service can be utilized from a Linux server. Q-Free's software is used in countries around the world for traffic management and toll collection. The wide-ranging geography of its use mean that Intrada ALPR is likely to be reliable and accurate for all types of license plates. Use of the Intrada suite is dependent on Q-Free granting an educational use license for this semester.

OpenALPR is another C++ library for use with both North American and European plates. This library relies upon two underlying technologies: OpenCV \cite{openCV} and Tesseract OCR (an OCR engine being developed and maintained by Google). The Tesseract OCR Tool is self-sufficient, and relies on included training data. \cite{patel2012} goes into greater detail how Tesseract extracts data from images. The open-source nature of this project make it appealing as it can be immediately integrated without needing to obtain an educational license first. Additionally, the last updates were pushed to the source repository in January. While it is desirable for the software we consider to be relatively up-to-date and maintained, this repository was started only four months ago (November, 2013). Use of such a young library may create a less stable solution than what is practical to work with throughout the course of the semester.

@@

Upon upload of the image to the ACMx server, a script will be invoked to process the image. The ACMx server will then run redundancy measures to ensure the image uploaded does contain a license plate. Upon a successful identification, the server will invoke the selected ALPR strategy to extract characters from the license plate image. Most of the image cropping will be done on the image acquisition Raspberry Pis to reduce the size of the data transmission, and any additional cropping and image manipulation required by the selected ALPR strategy will be done before invoking this script.

@@

Once the ACMx server has a plain text representation of the license plate characters, they will be stored in a MySQL database, along with a timestamp and lot id, where they can be retrieved and processed by the Web Application subsystem described in section 2.6.

\subsection{Web Application}
As mentioned in section 1.2, the ACMx group has already created a web application that will display the data collected from the sensors in the system. The site is currently accessible online at \url{http://acmxlabs.org/parking/}. Since the current site is essentially just a shell with no data, one main goal of this project is to make the main pages of the site fully functional. Another main goal is to create an administrator side to the site, where data can be viewed and updated by site administrators. The main focus of this subsystem will be to provide interfaces for both general users and administrators that make the data we collect both useful and easy to access. Since the ACMx team already has some protocols in place for collecting and displaying the data from the magnetometer Arduino Fios, the ACMx team will continue development on this interface, and our group will focus on other parts of the application, as described below. The web application itself will be written with PHP on the back end and HTML5/CSS3/JavaScript on the front end. The following subsections describe the various pieces of the site that we aim to complete during this project.

@@

\subsubsection{Validating and Adjusting Lot Capacity Data}
We expect that our first deployment of this system may not be 100\% accurate at detecting all vehicles that enter and exit lots. Therefore, as mentioned in section 2.4, we will be collecting periodic overhead images from a camera mounted on the base station Raspberry Pi. These images will serve as a visual representation of how many cars are actually in the parking lot at the time the image was taken. Administrators should be able to periodically compare this image to the collected data from the sensor network to make sure the numbers being computed by the network accurately represent the number of cars in the lot. The web interface will allow administrators to pull up the image for any lot with an overhead camera and see the numbers for that lot right next to the image. If any discrepancies are found, the administrator should be able to overwrite the count in the database to reflect reality. This interface will be extremely useful in helping to calibrate the system. For example, if the sensors are consistently reporting less cars in the lot than the image shows, we can fine-tune the collection process to make sure fewer cars slip through unnoticed. Future work in this area will include an automated version of this process, where a computer program will compare the image to the sensed data and make any necessary corrections.

@@

\subsection{Deployment}
The system will be deployed at the CTLM Upper and Lower parking lots at CSM. These lots were chosen because of their accessible location, unobstructed visibility from a campus structure, and wide use. Both parking lots have one entryway, each of which will be equipped with an automobile detection subsystem. The junction between the two lots is also of interest, because we want to know when a car moves from the upper lot to the lower lot, and vice versa. The locations of the entrances and junction are shown in Figure \ref{fig:deployment}. The locations for the three sensor nodes benefit from having a clear view of the sky for increased photovoltaic efficiency.
\begin{figure}
\begin{center}
\includegraphics[width=5in]{deployment}
\end{center}
\caption{Deployment Location}
\label{fig:deployment}
\end{figure}

The central base station, as described in section 2.4, will be placed in an upper window of the CTLM building, allowing it to take a clear overhead picture of at least one of the two parking lots. For the first deployment we may only have one overhead camera to start off with, although ideally we would have two - one for each lot. Several potential locations exist which allow for a clear line-of-sight to all automobile detection subsystems, one of which is detailed in Figure \ref{fig:deployment} The location of the base station must put it in range of the XBee Pros mounted on each sensor node, to allow data to be transferred from the detection subsystem to the base station. The locations being considered must also provide internet access to the base station Raspberry Pi through ethernet ports.

\section{Summary}
This project extends the work done by ACMx to create the first comprehensive parking management system for the Colorado School of Mines. The proposed expansions cover interfacing with the existing automobile detection network, acquiring images of license plates, processing those images, and presenting the resulting data in a web application.

Acquiring images of license plates will include experimentally testing the timing of taking the picture, as well as on-mote pre-processing of the image. By completing some processing of the image on the Raspberry Pi, we can both reduce the size of data we need to transmit (saving battery power), while also avoiding transmitting false positives (when no license plate is detected).

Image acquisition will come from several acquisition Raspberry Pis stationed at parking lot entrances. These images will be sent to a central base station, which will then upload the data to a server. Once on the server, additional image processing will be done to retrieve the characters composing the license plate. This data will be stored in a database for use in the web application.

In addition, the base station Raspberry Pi will sample images of the parking lots to allow administrators to correct the status of the system. This may happen when the acquisition motes unsuccessfully identify a car entering or exiting, and therefore have incorrect information regarding the parking lot's capacity.

The web application will serve to show the state of the parking lots, as well as offer a way for administrators to reconcile observed data (from base station images) and sensed data (from acquisition motes). Administrators will also be able to access license plate data from processed images to show proof-of-concept for this utility.

Once deployed, the SmartLots system, augmented with our additional license plate extraction subsystem and improvements to the web application, will offer the Mines campus an accessible way to manage campus parking lots. Faculty and students will be able to use this to find open parking and save time. Future applications could include interfacing with Parking Services to locate illegally parked vehicles or to gather data regarding campus parking needs.

@@

\bibitem{openCV} \emph{OpenCV} [Online] Available: \url{http://opencv.org/}
